---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
#To set path to the working directory, please paste your path below where 04cars.rda dataset is located and uncomment the code below
#setwd("D:/projects/R_Projects/Sandbox/Stats_model_assignments") 
#getwd()
```

### Question 1

Write a function confBand(x, y, conf=0.95) taking in a predictor vector (x1, . . . , xn) and a response vector y = (y1,...,yn) and return a plot with the points (x1,y1),...,(xn,yn), the least squares line, and the confidence band at level conf. Apply your function to hp and mpg from the 04cars dataset.

```{r}
#Loading the cars dataset
load("04cars.rda")
data = dat[,c(13,15)]
names(data) = c("hp", "mpg")

#defining the confBand function
confBand <- function(x,y, conf = 0.95){
  
  n = 100
  #Plotting the linear model
  linear_model = lm(y~x)
  p=1
  df_1 = p+1 #degree of freedom 1
  df_2=n-p-1  #degree of freedom 2
  fquartile = sqrt((p+1) * qf(conf, df_1,df_2))
  
  #Calculating y bar
  y_ = predict(linear_model, data.frame(x=x), se =  T)
  
  #Calculating the upper and lower confidence bounds
  cb_upper =  (y_$fit + ((fquartile)*y_$se.fit))
  cb_lower =  (y_$fit - ((fquartile)*y_$se.fit))
  

  plot(x,y, type = 'p')
  
  lines(x,cb_upper, type = 'l', col ="green")
  lines(x,cb_lower, type = 'l', col ="green")
  abline(linear_model, col = 'red', lwd = 2)
  legend("topright", legend=c("Linear model", "Conf band"),
       col=c("red", "green"), lty=1:2, cex=0.8, text.font=4, bg='001000')
  
  #return (list(upb = cb_upper, lwb = cb_lower, y_pred = y_$fit, pred_se = y_$se.fit, Fvalue = fquartile))
}

confBand(data$hp, data$mpg)



```

### Question 2

Let n = 100 and draw x1, . . . , xn ∼ Unif(0, 1), which stay fixed in what follows. Repeat the following experiment N = 1000 times.

• Generate yi = 1 + xi + εi, with εi i.i.d. N (0, 0.2). • Compute the 99% confidence band and record whether it contains the true line, or not.

Summarize the result of this numerical experiment by returning the proportion of times (out of N) that the confidence band contained the true line.

```{r}
x = runif(100, 0,1)
y_true = 1.000+x
contains_true_line = rep(NA, 1000)

# looping 1000 times
for (j in 1:1000){
  
  #calculating the error with N(0,0.2)
  e = rnorm(100, 0, 0.2)
  y = 1.000+x+e #generating y for every j
  
  CI = data.frame(lower = rep(NA,100),upper = rep(NA,100),contain_true_value = rep(NA,100))
  #fitting the linear model
  linear_model= lm(y~x)
  
  n = 100
  p=1
  df_1 = p+1
  df_2=n-p-1
  conf = 0.99
  F_quartile = sqrt((p+1) * qf(conf, df_1,df_2))
  pred_op = predict(linear_model, data.frame(hp=x), se =  T)
  

  CI$lower = (pred_op$fit - ((F_quartile)*pred_op$se.fit))
  CI$upper = (pred_op$fit + ((F_quartile)*pred_op$se.fit))
  CI$contain_true_value = (CI$lower <=y_true & CI$upper >= y_true) 
  
  #contains_true_line[j] = ((mean(CI$contain_true_value)) ==1)
  contains_true_line[j] = ((sum(CI$contain_true_value)) ==n)
}


#answer = (sum(contains_true_line)/1000)
answer = (sum(contains_true_line)/1000)

sprintf('Result of the experiment %f', answer)
```
